Neural models for target-based computer-assisted musical orchestration: a preliminary study

### 2020-05-16 paper/latex/paper.tex
final version before proof read

\usepackage{lineno}
%\linenumbers
\newcommand{\carmine}[1]{\textbf{\color{red} Carmine => #1}}

@@

%
\maketitle
%
\begin{abstract}
In this paper we will explore preliminary neural models for the task of target-based computer-assisted musical orchestration. After an introduction on computer-assisted orchestration, we will show how we model this  musical problem as a classification task. In this context, we will propose two deep learning models and we will show, first, how they perform as classifiers for musical instruments recognition by comparing them with specific baselines. We will then show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems. We will highlight, finally, benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

Computer-assisted orchestration is an important problem ... \carmine{Can you add more?}.

The approach studied in \cite{Cella18, Cella2020} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized. Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function $f$ and each generated combination of sounds is evaluated by using a specific metric. This method is currently considered state of the art and is implemented in the \emph{Orchidea} toolbox for assisted orchestration (\url{www.orch-idea.org}); we will describe this approach in more detail in the next sections.

In this paper, we try a different approach by experimenting with deep neural architectures. The main idea is to train a model to classify combinations of real instruments and use it then for orchestration. A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}. By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation. Thus, when an unknown target sound is given as input, the network will identify which are the best instruments-pitches pairs to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument-note pairs\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.

@@

In this paper we do preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}. We chose to use a CNN because of its success in audio classification \cite{Hershey16} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.

After training these models on varying combinations of 10 instruments, we used them for orchestration. More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system. We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means. For quantitative comparison, we computed the distance metric suggested in \cite{Cella2020} on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.

Next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results. Sections \ref{sec:orchestration} and \ref{sec:conclusions} will show, respectively, the results of orchestration experiments using the trained models and will present the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.

@@

@@

In this paper, we model assisted orchestration as a classification problem. The general methodology is as follows:
\begin{enumerate}
\item we train specific models to classify instrument-pitch pairs on sound combinations of a database of instrument notes, up to ten simultaneous instruments;
\item we then pick the best classifier and we feed into it an unknown sound to be classified;
\item since the output of the classifier will be made of instrument-pitch pairs, we use this information to synthesize an orchestration for the target sound;
\item finally, we evaluate the generated orchestration against state-of-the-art systems for computer-assisted orchestrations.
\end{enumerate}

@@

A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics. This is difficult, however, since to frame this problem as classification, we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample. For these reasons, our models have been trained on simplified versions of the original problem.

On one hand, the baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g. C, D\#, G, etc.) for a total of 25 classes.

On the other hand, for the neural models, we reframed the classification problem by using full pitch and trained them to recognize the pair instruments-pitch for a total of 424 classes (i.e. the neural classifiers would classify instruments and pitches with respective octaves). We want to point out, however, that this is still a simplified problem compared to assisted orchestration. Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note. To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models. When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}). By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.  Figure \ref{fig:method} illustrates the described ideas.

\begin{figure}
\includegraphics[scale=0.5]{figs/Acc_vs_N_CNN.png}
\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
\end{figure}

\begin{figure}
\center{
\includegraphics[scale=0.5]{figs/CNN_vs_ResNet.png}}
\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
\end{figure}

@@

To create the input data for training the classifiers, we used the \emph{TinySOL} database. TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM \cite{Cella2020b}. TinySOL contains 1,529 samples from 12 instruments. The instruments come from different orchestral families: strings, woodwinds, and brass. Each sample is one instrument playing a single note in the \emph{ordinario} playing style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff}, for example \texttt{Flute-C4-pp} or \texttt{Clarinet-D5-mf}. The instruments and ranges over which they were recorded is summarized in table \ref{tab:ranges}.

@@

The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{Stevens37} and can be considered to be an accurate representation of sound and musical signals. The Mel spectrogram is generated using an FFT hop length of 2048 samples (e.g. the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Instrument & Abbreviation & Range \\
\hline
Violin & Vn & G3-E7\\
Cello & Vc & C2-C6\\
Viola & Va & C3-C7\\
Trumpet C & TpC & F\#3-C6\\
Trombone & Tbn & A\#0;A\#1-C5\\
Oboe & Ob & A\#3-A6\\
Horn & Hn & G1-F5\\
Flute & Fl & B3-D7\\
Clarinet Bb & ClBb & D3-G6\\
Contrabass & Cb & E1-C5\\
Bass Tuba & BTb & F\#1-F4\\
Bassoon & Bn & A\#1-D\#5\\
\hline
\end{tabular}
\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
\end{center}
\end{table}

@@

In order to increase variability in our generated data, we also used two methods of data augmentation as described in \cite{Salamon17, hardwaj17}; more specifically, we used pitch shifting and partial feature dropout.

The former has been directly applied on the TinySOL samples each time the are selected to generate a new combination. We performed a small pitch shift by reading the samples with different sample rates: a small difference in sample rate will slightly modify the duration and the perceived pitch if played at the normal sample rate. In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.

@@

\subsection{Baseline}

In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we initially used various parametric classifiers. The classifiers we tested were support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN). We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}. More specifically, for SVM we used a non-linear RBF kernel and for RF we set the maximum depth of each tree to be 15.  All of the baseline experiments used 50,000 generated samples with a train-test split of 60/40. Each sample was made by a combination of one to ten instruments and was four seconds in length. 
In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a number of features that is more appropriate for the parametric classifiers. We found SVM to have the highest accuracy of the three classifiers across all experiments.

@@

This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments. However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased. Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with parametric classifiers.

@@

In the rest of this section, we will show the classification results of two neural models.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.4]{figs/CNN_LSTM_diagram.png}
\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
\end{center}
\end{figure*}

@@

During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks. However, the value of the loss function alone is difficult to interpret. For this reason we created a complementary function $f$ to be used only for evaluation. This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function

@@

\begin{figure}
\includegraphics[scale=0.5]{figs/Acc_vs_N_ResNet.png}
\caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
\end{figure}

@@

After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration. For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet. The total number of classes in this setting is 424, each one representing an instrument-note pair.

To orchestrate, a target sound is input to the model, and the 10 classes with the highest probability are extracted. These 10 classes are the instrument-note pairs that are most represented in the target, and can be from any combination of the 10 instruments.

@@

We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to Orchidea's solutions, the state-of-the-art program for computer-assisted orchestration. In order to have a more fair comparison, we did not allow Orchidea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution. This creates a more fair comparison since our model is unable to create sparse solutions.

Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch. For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution. For example, one of the samples of a bell had partials that represented a C\# minor chord, so we checked if the orchestration contained the notes of the chord. If a target had specific notes, we identified if the note or its partials were present. For example there is target of an Oboe playing an A4 and a Bassoon playing a C\#3, the solution from ResNet contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.

For quantitative evaluation, we used the distance metric defined in Eqn. \eqref{distance} to calculate differences in timbre between targets and solutions. The equation takes in the full FFT of the target and full FFT of the solution. Then for each bin of the FFT, it calculates the absolute difference between the values. The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.

In the first summation, $\lambda_1$ is multiplied by all the distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$. Similarly in the second summation, $\lambda_2$ is multiplied by all the distances when the solution provided more energy to a frequency than the target; it overshot the frequency. Therefore, the relation between $\lambda_1$ and $\lambda_2$ determines whether a solution is penalized more for undershooting or overshooting the target.

We calculated the distance between the target samples and our orchestrated solutions. We then orchestrated the same targets with OrchIdea and calculated the distance for OrchIdea's solutions. A comparison of these results is in Tab. \ref{orch_eval}.

\begin{equation}\label{distance}
d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
\end{equation}

\begin{equation}
\delta_{k1} = 
\begin{cases}
1, \text{if   } x_k \ge \tilde{x}_k \\
0, \text{otherwise}
\end{cases} 
\end{equation}

\begin{equation}
\delta_{k2} = 
\begin{cases}
1, \text{if   } x_k < \tilde{x}_k \\
0, \text{otherwise}
\end{cases}
\end{equation}

@@

Classification through deep learning seems promising for orchestration because of our model's ability to classify individual instruments and pitches out of dense combinations of samples. While our model is not able to outperform OrchIdea, it shows consistent results. When comparing our methods to OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution are not present in our model. Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.

We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets. Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.

Future steps in this project include implementing various methods that are present in OrchIdea.

Our current model orchestrates all targets using the same number of samples, and this does not take into account the density of different targets. The solution to this is to allow sparse solutions in which the model decides how many samples should be used to best represent the target. This allows a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.

Partial filtering is a method that would aid our model in orchestrating the harmonics of a target. The dominant harmonic partials of the target are identified, and the search space is limited to only include samples of those pitches. For example, if the target is a recording of an instrument playing a C4, then the partials identified may be C4, C5, G5, and E6. The model would then only consider samples of these pitches to be used in the solution. This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.

@@

\bibitem{Benetos07}
Emmanouil Benetos, Margarita Kotti, and Constantine Kotropoulos.
\newblock Large scale musical instrument identification.
\newblock 2007.

@@

\bibitem{Bian19}
Wenhao Bian, Jie Wang, Bojin Zhuang, Shaojun~Wang Jiankui~Yang, and Jing Xiao.
\newblock Audio-based music classification with densenet and data augmentation.
\newblock ArXiv (https://arxiv.org/abs/1906.11620v1), 2019.

\bibitem{Cella18}
Carmine~Emanuele Cella and Philippe Esling.
\newblock Open-Source Modular Toolbox for Computer-Aided Orchestration.
\newblock Timbre conference, 2018.

\bibitem{Cella19}
Carmine~Emanuele Cella, Jon Gillick, and David Bamman.
\newblock Estimating unobserved audio features for targed-based orchestration.
\newblock ISMIR, 2019.

\bibitem{Cella2020}
Carmine~Emanuele Cella.
\newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
\newblock in preparation; available on request from the author, 2020.

@@

\bibitem{Chetry05}
Nicolas Chetry, Mike Davies, and Mark Sandler.
\newblock Musical instrument identification using lsf and k-means.
\newblock 2005.

\bibitem{He15}
Kaiming He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
\newblock Deep residual learning for image recognition.
\newblock 2015.

>>>>

\usepackage{lineno}
%\linenumbers
\newcommand{\carmine}[1]{\textbf{\color{red} Carmine => #1}}
\renewcommand{\baselinestretch}{.98}

@@

%
\maketitle
%
\begin{abstract}
In this paper we will do a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration. We will show how it possible to model this  musical problem as a classification task and we will propose two deep learning models. We will show, first, how they perform as classifiers for musical instruments recognition by comparing them with specific baselines. We will then show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems. Finally, we will highlight benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

The development of computational tools to assist and inspire the musical composition process constitutes an import research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}. Within CAC, target-based computer-assisted orchestration is an important example how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.

Target-based computer-assisted orchestration can be thought of as the process of searching for the best combinations of sounds in a database of instrumental samples, to match a target sound under a specific similarity metric and specific sets of constraints. A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound. 
Valid orchestral scores may contain several instruments sounding simultaneously, selected out of a large number of possible combinations of sounds from the database.

\begin{figure}
\includegraphics[scale=0.17]{figs/method.png}
\caption{An overview of the proposed method for assisted orchestration with neural models. Instruments and pitches are determined as peaks of the output probability distribution, while the dynamics are computed by quantizing the probabilities. \label{fig:method}}
\end{figure}

The approach studied in \cite{Carpentier2010} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized. Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function and each generated combination of sounds is evaluated by using a specific metric. This method has been substantially improved in \cite{Cella18, Cella2020} and is implemented in the \emph{Orchidea} toolbox for assisted orchestration (\url{www.orch-idea.org}), currently considered the state-of-the-art system for assisted orchestration.

In this paper, we try a different approach to this problem by experimenting with deep neural architectures. The main idea is to train a model to classify combinations of real instruments and use it then for orchestration. A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}. By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation. Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes.

%\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.

@@

In this paper we do preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}. We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.

After training these models on varying combinations of 10 instruments, we used them for orchestration. More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system. We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means. For quantitative comparison, we computed a specific distance metric on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.

Next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results. Sections \ref{sec:orchestration} and \ref{sec:conclusions} will present, respectively, the results of orchestration experiments using the trained models and the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.22]{figs/CNN_LSTM_diagram.png}
\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
\end{center}
\end{figure}

@@

In this paper, we model assisted orchestration as a classification problem. The general methodology is as follows:
\begin{enumerate}
\item we train specific models to classify the instruments present in combinations of sounds from a database of instrument notes, up to ten simultaneous instruments;
\item we then pick the best classifier and we feed into it an unknown sound to be classified;
\item since the output of the classifier will be in the form of the probability that specific instruments are present in the sound, we use this information to synthesize an orchestration for the target sound;
\item finally, we evaluate the generated orchestration against state-of-the-art systems for computer-assisted orchestrations.
\end{enumerate}

\begin{figure}
\center{
\includegraphics[scale=0.4]{figs/CNN_vs_ResNet.png}}
\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
\end{figure}

@@

A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample. For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.

In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.

Three baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g. C, D\#, G, etc.) for a total of 25 classes.

Two neural classifiers, instead, have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e. the neural classifiers would classify instruments and pitches with respective octaves). We want to point out, however, that this is still a simplified problem compared to assisted orchestration. Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note. To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models. When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}). By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.  Figure \ref{fig:method} illustrates the described ideas.

@@

To create the input data for training the classifiers, we used the \emph{TinySOL} database. TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM \cite{Cella2020b}. TinySOL contains 1,529 samples from 12 instruments. The instruments come from different orchestral families: strings, woodwinds, and brass. Each sample is one instrument playing a single note in the \emph{ordinario} playing style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff} (for example \texttt{Flute-C4-pp} or \texttt{Clarinet-D5-mf}).

%The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.

@@

The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an accurate representation of sound and musical signals. The Mel spectrogram is generated using an FFT hop length of 2048 samples (e.g. the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.

%\begin{table}
%\begin{center}
%\begin{tabular}{|c|c|c|}
%\hline
%\textbf{Instrument} & \textbf{Abbreviation} & \textbf{Range} \\
%\hline
%Violin & Vn & G3-E7\\
%Cello & Vc & C2-C6\\
%Viola & Va & C3-C7\\
%Trumpet C & TpC & F\#3-C6\\
%Trombone & Tbn & A\#0;A\#1-C5\\
%Oboe & Ob & A\#3-A6\\
%Horn & Hn & G1-F5\\
%Flute & Fl & B3-D7\\
%Clarinet Bb & ClBb & D3-G6\\
%Contrabass & Cb & E1-C5\\
%Bass Tuba & BTb & F\#1-F4\\
%Bassoon & Bn & A\#1-D\#5\\
%\hline
%\end{tabular}
%\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
%\end{center}
%\end{table}

@@

In order to increase variability in the generated data for the neural models, we also used two methods of data augmentation as described in \cite{Salamon17, Bhardwaj17}; more specifically, we used pitch shifting and partial feature dropout.

The former has been directly applied on the TinySOL samples each time they are selected to generate a new combination. We performed a small pitch shift by reading the samples with different sample rates: a small difference in sample rate will slightly modify the duration and the perceived pitch if played at the normal sample rate. In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.

@@

\subsection{Baseline}
The baseline classifiers that we tested are support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN). We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}. More specifically, for SVM we used a non-linear RBF kernel and for RF we set the maximum depth of each tree to be 15.  All of the baseline experiments used 50,000 generated samples with a train-test split of 60/40. Each sample was made by a combination of one to ten instruments and was four seconds in length. 
In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a number of features that is more appropriate for the parametric classifiers. We found SVM to have the highest accuracy of the three classifiers across all experiments.

@@

This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments. However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased. Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments.

@@

In the rest of this section, we will show the classification results of the two neural models.

@@

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.4]{figs/Acc_vs_N_CNN.png}
\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
\end{center}
\end{figure}

During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks. However, the value of the loss function alone is difficult to interpret. For this reason we created a complementary function $f$ to be used for evaluation only. This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.40]{figs/Acc_vs_N_ResNet.png}
\caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
\end{center}
\end{figure}

@@

After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration. For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet. The total number of classes in this setting is 424, each one representing an instrument-pitch pair.

To orchestrate, a target sound is input to the model, and the 10 classes with the highest probability are extracted. These 10 classes are the instrument-pitch pairs that are most represented in the target, and can be from any combination of the 10 instruments.

@@

We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to the solutions generated by Orchidea, the state-of-the-art system for computer-assisted orchestration.

Orchidea implements many advanced features that are not supported by our models. For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles. It is also able to reduce the search space by applying harmonic analysis on the target sound.
The dominant harmonic partials of the target are identified, and the search space is limited to only include samples of those pitches. For example, if the target is a recording of an instrument playing a C4, then the partials identified may be C4, C5, G5, and E6. The model would then only consider samples of these pitches to be used in the solution. This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.
Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this propriety is called \emph{sparsity} of the solution).

In order to have a more fair comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution. This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.

Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch. For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution. For example, one of the samples of a bell had partials that loosely represented a C minor chord, so we checked if the orchestration contained the notes of the chord. If a target had specific notes, we identified if the note or its partials were present. For example there is target of an Oboe playing an A4 and a Bassoon playing a C\#3, the solution from ResNet contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.

For quantitative evaluation, we used the distance metric defined in Eqn. \eqref{distance} to calculate differences in timbre between targets and solutions. This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization. The equation takes in the full FFT of the target $x$ and full FFT of the solution $\hat{x}$. Then for each bin $k$ of the FFT, it calculates the absolute difference between the values. The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.

In the first summation, $\lambda_1$ is multiplied by all the distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$. Similarly in the second summation, $\lambda_2$ is multiplied by all the distances when the solution provided more energy to a frequency than the target. Therefore, the relation between $\lambda_1$ and $\lambda_2$ determines whether a solution is penalized more for undershooting or overshooting the target.

We calculated the distance between the target samples and our orchestrated solutions. We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions. A comparison of these results is in Tab. \ref{orch_eval}.

\begin{equation}\label{distance}
d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
\end{equation}
where 
$
\delta_{k1} = 
\begin{cases}
1, \text{if   } x_k \ge \tilde{x}_k \\
0, \text{otherwise}
\end{cases} 
$, $
\delta_{k2} = 
\begin{cases}
1, \text{if   } x_k < \tilde{x}_k \\
0, \text{otherwise}
\end{cases}.
$\\

While our model is not able to outperform Orchidea, it shows consistent results. We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets. Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.

@@

Target-based computer-assisted orchestration through deep learning models seems a promising path, thanks to the ability of deep networks to classify individual instruments and pitches out of dense combinations of samples. This work, however, represents only a preliminary study of the potential of these methods for the task of assisted orchestration.

The first natural extension would be to support sparsity in our models. Our current models, indeed, orchestrate all targets using a constant number of instruments and it is not able to drop specific instruments from the solution. This does not take into account the density of different targets. By using sparse solutions, in which the model decides how many samples should be used to best represent the target, would allow a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds. This would generate more meaningful orchestrations that would compare more favourably to the state of the art.

Another important extension would be to create a more powerful embedding spaces for the target and combinations. In \cite{Gillick19} the authors propose to use LSTM-based models to predict the embedding features for the combinations used during the optimisation process in assisted orchestration. We believe that by combining their prediction model with our classification models could generate more faithful representations and improve the overall quality of generated orchestrations.

@@

\bibitem{FerVic2013} J.D. Fern\'andez and F. Vico. {AI Methods in Algorithmic Composition: A Comprehensive Survey}. Journal of Artificial Intelligence Research, vol.48, pp. 513--582, 2013.

\bibitem{Ari2005} C. Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}. In Proceedings of the International Computer Music Conference, 2005.

\bibitem{Maresz2003}
Yan Maresz, {On Computer-Assisted Orchestration}. Contemporary Music Review, vol. 32, n. 1, pg. 99-109, Routledge, 2013.

\bibitem{Carpentier2010}
Gr\'egoire Carpentier, Damien Tardieu, Jonathan Harvey, G\'erard Assayag and Emmanuel Saint-James. {Predicting Timbre Features of Instrument Sound Combinations: Application to Automatic Orchestration}, Journal of New Music Research, vol. 39, n. 1, pg. 47-61, Routlegde, 2010.

\bibitem{Benetos07}
Emmanouil Benetos, Margarita Kotti, and Constantine Kotropoulos.
\newblock Large scale musical instrument identification.
\newblock Sound and Music Computing Conference, Greece, 2007.

@@

\bibitem{Bian19}
Wenhao Bian, Jie Wang, Bojin Zhuang, Shaojun~Wang Jiankui~Yang, and Jing Xiao.
\newblock Audio-based music classification with densenet and data augmentation.
\newblock arXiv preprint, https://arxiv.org/abs/1906.11620v1, 2019.

\bibitem{Cella2020}
Carmine~Emanuele Cella.
\newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
\newblock In preparation; available on request from the author, 2020.

@@

\bibitem{Chetry05}
Nicolas Chetry, Mike Davies, and Mark Sandler.
\newblock Musical instrument identification using LSF and K-means.
\newblock Audio Engineering Society Convention 118, 2005.

\bibitem{He15}
Kaiming He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
\newblock Deep residual learning for image recognition.
\newblock arXiv preprint, https://arxiv.org/abs/1512.03385, 2015.

\bibitem{Hershey17}
Shawn Hershey, Sourish Chaudhuri, Daniel P.~W. Ellis, Jort~F. Gemmeke, Aren
  Jansen, R.~Channing Moore, Manoj Plakal, Devin Platt, Rif~A. Saurous, Bryan
  Seybold, Malcolm Slaney, Ron~J. Weiss, and Kevin Wilson.
\newblock CNN architectures for large-scale audio classification.
\newblock IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP), 2017.
